Nitric-oxide-feedback to ciliary photoreceptor cells orchestrates UV avoidance in zooplankton

### 2022-05-28 text/Manuscript.Rmd

---
title: "Nitric-oxide-feedback to ciliary photoreceptor cells orchestrates UV avoidance in zooplankton"
author: "Kei Jokura, Nobuo Ueda, Martin Gühmann, Gáspár Jékely"
date: "*Living Systems Institute, University of Exeter, Stocker road, Exeter, EX4 4QD, UK*"
output:
  pdf_document: 
      latex_engine: xelatex
      template: "default"
  html_document: 
    theme: lumen
geometry: left=0.8in,right=0.8in,top=0.8in,bottom=0.5in
mainfont: Arial
documentclass: article
classoption:
  - onecolumn
linestretch: 1.2
fontsize: 10pt
bibliography: references.bib
csl: elife.csl
---

@@

![Figure 1. Legend goes here](../figures/Figure1.png){width="100%"}

@@

![Figure 2. Legend goes here](../figures/Figure2.png){width="100%"}

@@

![Figure 3. Legend goes here](../figures/Figure3.png){width="100%"}

@@

![Figure 4. Legend goes here](../figures/Figure4.png){width="100%"}

@@

![Figure 6. Legend goes here](../figures/Figure6.png){width="100%"}

>>>>

---
title: "Nitric-oxide-feedback to ciliary photoreceptor cells orchestrates UV avoidance in zooplankton"
author: "Kei Jokura, Nobuo Ueda, Martin Gühmann, Gáspár Jékely"
date: "*Living Systems Institute, University of Exeter, Stocker road, Exeter, EX4 4QD, UK*"
output:
  html_document: 
    theme: lumen
  pdf_document: 
      latex_engine: xelatex
      template: "default"
geometry: left=0.8in,right=0.8in,top=0.8in,bottom=0.5in
mainfont: Arial
documentclass: article
classoption:
  - onecolumn
linestretch: 1.2
fontsize: 10pt
bibliography: references.bib
csl: elife.csl
---

@@

![**Figure 1. Identification of NOS-expressing interneurons (INNOS) within the cPRC circuit.** (A and B) Immunostaining with anti-HA antibody against HA tagged palmitoylated-tdTomato expressed under the upstream of the start site of NOS (magenta). Co-staining with acetylated α-tubulin antibody as marker for cilia and axonal scaffold (ac α-tub: green). INNOSs with the entire cell body and axon labelled are shown with white arrowheads and with only the cell body labelled are shown with black arrowheads. Views of larvae at 2-day-old (A) and 3-day-old (B) from the anterior side, respectively. (C) Three types of neurons (cPRC, INNOS and INRGWa) in the cPRC circuit reconstructed in the whole-body transmission electron microscopy (ssTEM) dataset of a 3-day-old larva. The INNOS on the ventral side labelled with (A) and (B) are shown in purple. Other neurons are shown in light grey. Axons and dendrites appear as lines and cell body positions are represented by spheres. (D and E) Expression analysis of the NOS gene (magenta) using in situ HCR. Anterior view of the larva at 2-day-old. (D) Co-stained image using acetylated α-tubulin antibody (ac α-tub: green). (E) Co-expression analysis image with the RYamide proneuropeptide gene (RYa-pNP: green), which is highly expressed in INNOS. White arrowheads indicate co-expressed INNOS. Black arrowheads indicate larval eyes. White dashed lines indicate the outer frame of the larva. (F) Wiring diagram of the cPRC circuit. Hexagons represent neuronal groups (numbers indicated in square brackets), arrows represent synaptic connections (numbers indicated in number of synapses); NOS and RGW indicate INNOS and INRGW-dc1 respectively. Ciliomotor neuron shows two serotonergic neurons (Ser-h1).](../figures/Figure1.png){width="100%"}

@@

![**Figure 2. Larvae with inhibited NOS have shorter ciliary arrest times and a delayed response to UV-avoidance.** (A) Percentage of time in which cilia are closed in DMSO-treated larvae (control) vs. larvae exposed to NOS inhibitor. (B) Top: The domain structure of NOS. Bottom: Close-up region showing the genomic locus of NOS gene and the wild-type sequence (WT) targeted by CRISPR/Cas9. The generated mutants (NOSΔ11, NOSΔ23) are also shown. Pink indicates target sites. Yellow shows PAM sequences, red shows stop codons. (C and D) Vertical swimming in wild type and mutant larvae at 2-day-old (C) and 3-day-old (D) stimulated with blue (480 nm) and UV (395 nm) light from above. the data are shown in 30 s bins.](../figures/Figure2.png){width="100%"}

@@

![**Figure 3. cPRC-specific expression of NIT-GC1 localises in close proximity to NOS in the ANS region.** (A) Phylogenetic tree of guanylate cyclase by maximum likelihood (ML). Guanylate cyclase-coupled receptor and soluble guanylate cyclase as outgroups. Guanylate cyclases with NIT domains are found in most animal phyla except Porifera, Ctenophora, Urochordata and Chordata. Tree robustness was tested with 1000 replicates of ultrafast bootstrap. Species names are as follows: Ap (Aiptasia pallida), Bb (Branchiostoma belcheri), Cv (Crassostrea virnigica), Dm (Drosophila melanogaster), Dp (Daphnia pulex), Hh (Hoilungia hongkongensis), Lg (Lottia gigantea), Pc (Priapulus caudatus), Pd (Platynereis dumerilii), Sk (Saccoglossus kowalevskii), Sp (Strongylocentrotus purpuratus), Ta (Trichoplax adhaerens), Tc (Tribolium castaneum). (B-D) Expression analysis of the NIT-GC1 gene (magenta) using in situ HCR. White arrowheads indicate cells expressing NIT-GC1. (B) Co-staining image using acetylated α-tubulin antibody (ac α-tub: green). (C) Co-expression analysis image of the c-opsin1 gene (green). Black arrowheads indicate cells expressing c-opsin1 differently from cPRC. (D) Co-expression analysis image of the MLD-pedal2 amide proneuropeptide gene (MLD-pNP: green). Black arrowheads indicate cells different from cPRC. The white dashed lines in (C) and (D) indicate the outer frame of the larvae. (E, F) Localisation analysis using NIT-GC1 and NIT-GC2 antibodies. Green shows co-staining with acetylated α-tubulin antibody (ac α-tub). (G) Co-localisation analysis using NIT-GC1 and NOS antibodies. (B-G) Anterior view of the 2-day-old larvae.](../figures/Figure3.png){width="100%"}

@@

![**Figure 4. NIT-GC1 is induced by NO to produce cGMP.** (A) Top: The domain structure of NIT-GC1. Grey squares indicate the transmembrane domain. Bottom shows a schematic of the NIT-GC1 assay; COS7 cells were used to express NIT-GC1 and the cGMP indicator Green cGull, and the NO donor SNAP and DMSO as control. Green cGull shows cGMP-dependent green fluorescence. (B) The intensity changes in cGMP fluorescence over time for the three conditions are shown. Gray traces show the optical responses to individual intensities (n = 6). The individual intensities are normalized (ΔF/F0). The dark line shows the mean value and the band the standard error range of the mean. The respective solution was acted on 2 min after the start of imaging (grey bars). (C) The mean intensity (Fmean/F0) after acting on the solution. Multiple comparisons were made using the Tukey honestly significant difference test.](../figures/Figure4.png){width="100%"}

@@

![**Figure 6. Gene expressions and signalling mechanisms between cPRC, INNOS and INRGWa.** (A) Dot plot of genes (columns) expressed in three types of cells (rows) in the cPRC circuit using single cell RNA-Seq. The size of the dots is expressed in proportion to the percentage of cells expressing that gene relative to all cells. The colours represent the normal logarithm of the number of transcripts in the cells expressing the gene. (B) Schematic diagram of the signalling pathway of the cPRC circuit, focusing on the NO feedback.](../figures/Figure6.png){width="100%"}
